Optimal updating magnitude in adaptive flat-distribution-sampling simulations

### 2017-09-09 doc/invt.tex
small edits to invt.tex

Free energy calculation\cite{frenkel, newman} is a central theme
in computational physics and chemistry
that can provide insight into an array of phenomena not easily studied
with regular simulation methods.
\note{``Regular simulation methods'' means MC and MD,
and they are more pertinent to ``computational physics and chemistry''
than ``traditional experiments.''}%
%
Given a system,
often the task is to compute
a distribution, $p(z)$,
along a collective variable, $z$, by sampling via either
Monte Carlo (MC)\cite{frenkel, newman, landau_binder}
or molecular dynamics (MD)\cite{frenkel, karplus2002}
simulations.
%
The distribution is related to
the dimensionless free energy, or potential of mean force (PMF),
along the collective variable, $z$,
as $-\ln p(z)$.
%
However, reliable estimation of the PMF is often difficult
when the collective variable is energetically restricted to local regions
of the complicated, multimodal free energy surface.
%
Thus,
to overcome these energetic barriers in sampling,
%and to capture the global shape of the free energy surface
an effective strategy is to introduce a bias potential that
offsets the PMF
such that the distribution along $z$ is
flat.\cite{mezei1987, berg1992, *lee1993,
wang2001, wang2001pre,
huber1994,
*laio2002, *laio2008, *barducci2011, *sutto2012}
%
\note{The flat-distribution sampling
allows one to more accurately capture the
global shape of the free energy surface
by improving the sampling around the rarely visited barrier region
and accelerating transitions among free energy basins.}

Many efficient flat-distribution-sampling (FDS) techniques
based on adaptive construction of the bias potential
have been introduced,
including the Wang-Landau (WL) algorithm\cite{
  wang2001, wang2001pre}
and metadynamics\cite{huber1994,
  *laio2002, *laio2008, *barducci2011, *sutto2012, micheletti2004}
among others\cite{yan2004, kim2006, *kim2007, kim2010, junghans2014,
  langfeld2012, pellegrini2014,
  maragliano2006, *abrams2008,
  zheng2010}.
%
These techniques regularly update the bias potential
to discourage future visits to previously sampled configurations
by incremental elevation of the bias potential.
%
A key difference lies
in the updating window function,
which specifies
a neighborhood around the current value of
$z$ on the free energy surface
%where the updating should occur
as well as the relative updating magnitude.
%
The window function often
takes the form of a discrete
$\delta$-function (a boxcar function one histogram bin wide)
in the WL algorithm,
or that of a Gaussian
in metadynamics.\cite{junghans2014}
%
We will refer to the two updating schemes
as the single-bin
and Gaussian updating schemes, respectively.
%
While the former is popular for MC simulations,\cite{wang2001,
  wang2001pre, kim2006, *kim2007}
the latter is commonly used for MD simulations
as it provides a smoother bias potential
that can be readily differentiated
to derive the bias force.\cite{huber1994,
  *laio2002, *laio2008, *barducci2011, *sutto2012, junghans2014}
%
In both cases,
regular updates to the bias potential
disrupts the underlying equilibrium
sampling,\cite{zhou2005, morozov2007, zhou2008}
and the convergence of the bias potential
can benefit from a gradually decreasing
updating magnitude of the bias potential.

The schedule of reducing the updating magnitude
affects the precision and accuracy
of the final bias potential,
and hence that of the PMF.\cite{
belardinelli2007, *belardinelli2007jcp, *belardinelli2008, *belardinelli2016,
liang2007, min2007,
morozov2007, morozov2009, zhou2008,
komura2012, *caparica2012, *caparica2014,
barducci2008, dickson2011, dama2014, poulain2006}
%
While an overly slow reduction of the updating magnitude
is ineffective in restricting the variance of the bias potential,
a too rapid reduction often
converges the bias potential to a wrong value.
%
For the single-bin case, it is known that
the optimal schedule is given by the
inverse time\cite{
belardinelli2007, *belardinelli2007jcp, *belardinelli2008, *belardinelli2016,
liang2007,
morozov2007, morozov2009, zhou2008}
or
the inverse number of updating steps.

@@

We will then show by comparing the errors
of different updating schemes
that the single-bin updating scheme
is optimal for very long simulations.
%without a priori assumption of the smoothness of the PMF.
%
A similar but partial asymptotic optimality applies to
a class of bandpass updating schemes
that flatten only
a few long-range distribution modes.
%
Compared with the Gaussian updating scheme,
bandpass updating schemes have the advantage of allowing
simpler control of the updating magnitude
for optimal convergence.

@@

From Eqs.
\eqref{eq:vt_diffeq} and \eqref{eq:sh_ave},
we get % a set of decoupled equations
%
\begin{equation}
  \dot v_{*i}(t)
  %=
  %\alpha(t) \, \left[ \langle f_{*i}(t) \rangle + \xi_{*i}(t) \right]
  \approx
  -\alpha(t) \, \left[ v_{*i}(t) - \xi_{*i}(t) \right]
  .
  %\notag
  \label{eq:dxdt_singlebin}
\end{equation}
%
Without the random part, $\xi_{*i}(t)$,
the deviation of the bias potential, $v_{*i}(t)$,
would tend to $0$ at long times, reaching convergence.
%
The formal solution of Eq. \eqref{eq:dxdt_singlebin} is
%
\begin{equation}
  v_{*i}(T)
  =
  v_{*i}(0) \, e^{-q(T)}
  +
  \int_0^T
    \dot{\vartheta}\bigl( q(t) \bigr) \, \xi_{*i}(t) \, dt,
  \label{eq:xt_solution}
\end{equation}
%
where
%
$q(t) \equiv \int_0^t \alpha(t') \, dt'$,
%
and
%
\begin{align}
  \vartheta(q')
  &\equiv
  e^{q' - q(T)}.
  \label{eq:theta_def}
\end{align}

@@

From Eqs. \eqref{eq:error_def} and \eqref{eq:xt_solution},
we can write the total error at the end of period $T$
as a sum of a residual error, $E_R$ (from the decay of the initial error)
and an asymptotic error, $E_A$:
%
\begin{align}
  \Err(T)
  =
  \Err_R(T) + \Err_A(T)
  ,
  \label{eq:error_tot}
\end{align}
%
where
\begin{align}
  \Err_R(T)
  &= \Err(0) \, e^{-2 \, q(T)}
  ,
  \label{eq:ER_sbin}
  \\
  \Err_A(T)
  &=
  \sum_{i=1}^n \rho_i
    \int_0^T\!\!\!\int_0^T
    \!\!
    \langle
      \xi_{*i}(t) \xi_{*i}(t')
    \rangle
    \dot \vartheta\bigl( q(t) \bigr)
    \dot \vartheta\bigl( q(t') \bigr)
    \, dt dt'
  \notag \\
  &= \Gamma \, \int_0^T \dot \vartheta^2\bigl( q(t) \bigr) \, dt
  ,
  \label{eq:EA_sbin}
\end{align}
with
%$\Gamma \equiv \sum_{i=1}^n \rho_i \, G_i$.
$\Gamma \equiv \sum_{i=1}^n \rho_i \, G_i
\approx \sum_{i=1}^n \rho_i \int_{-\infty}^{\infty} \langle \xi_{*i}(t) \xi_{*i}(0) \rangle \, dt$.
%\begin{equation}
%  \Gamma \equiv \sum_{i=1}^n \rho_i \, G_i
%  =
%  \sum_{i=1}^n \rho_i \int_0^T
%    \langle \xi_{*i}(t) \xi_{*i}(0) \rangle \, dt
%  .
%  \notag%\label{eq:Gamma_from_G}
%\end{equation}

The error depends implicitly on the curve, $\alpha(t)$,
or, equivalently, the curve, $q(t)$.
%via $v_{*i}(T)$.
%
%If $\alpha(t)$ is a constant, $a_0$,
%so that $q(t) = a_0 \, t$,
%then after a long period,
%the error approach an equilibrium value
%%
%\begin{equation}
%  \Err_\mathrm{equil.}
%  = \Err_A = \Gamma \int_0^\infty e^{-2\,q'} \, a_0 \, d q'
%  = \frac{1}{2} a_0 \Gamma
%  .
%  \label{eq:error_equil_sbin}
%\end{equation}
%
Our aim is to find the $q(t)$
that minimizes this error.
%
It is convenient to
vary the curve, $q(t)$ ($0 < t < T$),
under a fixed value of $q(T)$,
since it allows us to focus on
the asymptotic error, $E_A(T)$.
%
The contribution of the residual error
will be considered later
by adjusting the value of $q(T)$.
%
Note that the asymptotic error
has the form of an action of a free particle
at $\vartheta\bigl( q(t) \bigr)$,
with the endpoints fixed at
$\vartheta\bigl( q(0) \bigr)  = e^{- q(T)}$
and
$\vartheta\bigl( q(T) \bigr) = 1$.
%
Thus, the velocity should be a constant
%Since the residual error
%is fixed during the process,
%the asymptotic error,
%resembling the action of a free particle,
%demands a constant velocity
%
\begin{equation}
  \frac{d}{dt} \vartheta\bigl( q(t) \bigr) = c
  .
\label{eq:dthetadt_const}
\end{equation}
%
Then, by using Eq. \eqref{eq:theta_def}, we have
%
\begin{equation}
  e^{ q(t) - q(T) }
  =
  c \, (t + t_0)
  =
  \frac{ t + t_0 }
       { T + t_0 }
  ,
\label{eq:expqt}
\end{equation}
%
where $t_0$ and $c$ are two constants,
and the latter has been determined
from the $t = T$ case
in the second step.
%
Taking the logarithm,
and differentiating both sides
with respect to $t$,
we get the inverse-time schedule
%
\begin{equation}
  \alpha(t) = \frac{ 1 }{ t + t_0 }
  ,
\label{eq:alpha_invt1}
\end{equation}
%
with Eq. \eqref{eq:alpha_invt} being
the special case of $t_0 = 0$.
%

@@

Equation \eqref{eq:vt_diffeq_mbin} shows that
in the absence of the fluctuating term, $\tilde \xi_k(t)$,
the deviation of the bias potential would tend to zero
at a rate of $\alpha(t) \, \lambda_k$.
%
Thus, the updating scheme is convergent
only if no eigenvalue, $\lambda_k$, is negative
for $k > 0$.
%
%With a set of heterogeneous eigenvalues,
%the errors of the eigenmodes are reduced at different rates.
%Thus, the optimization problem of the final error
%becomes more complicated and it often
%involves a compromise among eigenmodes.
%
The formal solution is given by
\begin{equation}
  \tilde v_k(T)
  =
  \tilde v_k(0) \, e^{-\lambda_k \, q(T)}
  +
  \int_0^T
    \dot{\vartheta}_k\bigl( q(t) \bigr) \, \tilde\xi_k(t) \, dt
  ,
  \notag
  %\label{eq:xt_solution}
\end{equation}
where
\begin{equation}
  \vartheta_k(q') \equiv e^{\lambda_k \, [q' - q(T)]}
  .
  \label{eq:uk_def}
\end{equation}

@@

By Eq. \eqref{eq:parseval},
we can rewrite the error defined in Eq. \eqref{eq:error_def} as
a sum of the errors in individual modes\footnote{If
the error weighting distribution differs from $\rho_i$
in Eq. \eqref{eq:error_def},
the error would contain quadratic cross terms among different eigenmodes.
%
While the solution for the optimal schedule retains the same form,
the mass function defined in Eq. \eqref{eq:mass_func}
will include the corresponding cross terms.
}
%$\Err(T) = \sum_{k = 1}^{n - 1} \left\langle {\tilde v}_k^2(T) \right\rangle$.
%
\begin{align}
  \Err(T)
  =
  \sum_{k = 1}^{n - 1}
    \left\langle
      {\tilde v}_k^2(T)
    \right\rangle
  .
  \notag
\end{align}
%
As in Eq. \eqref{eq:error_tot},
we can decompose the error as the sum
of the residual error and the asymptotic error, with
\begin{align}
  \Err_R(T)
  &=
  \sum_{k = 1}^{n-1}
    \left\langle
      {\tilde v}_k^2(0)
    \right\rangle \,
    e^{ - 2 \, \lambda_k  \, q(T) }
  ,
  \label{eq:error_res}
  \\
  \Err_A(T)
  &=
  \sum_{k=1}^{n-1}
  \int_0^T\!\!\!\int_0^T
  \!\!
  \langle
    \tilde\xi_k(t)
    \tilde\xi_k(t')
  \rangle
  \vartheta_k\bigl(q(t)\bigr)
  \vartheta_k\bigl(q(t')\bigr)
  \, dtdt'
  \notag \\
  &=
  \int_0^T
  \sum_{k = 1}^{n-1}
  \Gamma_k \, \dot \vartheta_k^2\bigl( q(t) \bigr) \, dt
  ,
  \label{eq:error_asym}
\end{align}
%
where
%
%$\Gamma_k$ is computed from
%the autocorrelation function
%of the histogram fluctuation,
%$\bigl\langle {\tilde \xi}_k(0)
%\, {\tilde \xi}_k(t) \bigr\rangle$,
%as:
%
\begin{equation}
  \Gamma_k
  \approx
  \int_{-\infty}^\infty
  \bigl\langle
    \tilde\xi_k(t) \, \tilde\xi_k(0)
  \bigr\rangle
  \, dt
  \to
  \sum_{t = -\infty}^\infty
  \bigl\langle
    \tilde\xi_k(t) \, \tilde\xi_k(0)
  \bigr\rangle
  ,
  \label{eq:Gamma_sum}
\end{equation}
%
with the second expression being
the counterpart in the discrete-time setting.
%
Equation \eqref{eq:Gamma_sum} represents
a first approximation for modeling
a general fluctuation, ${\tilde \xi}_k(t)$,
as an equivalent white noise.
In this way,
the underlying sampling process
affects the error only through the
few numbers, $\Gamma_k$'s.
%
Note that by Eq. \eqref{eq:parseval}, we have
$\sum_{k=1}^{n-1} \bigl\langle {\tilde \xi}_k(t) \, {\tilde \xi}_k(0) \bigr\rangle
= \sum_{i=1}^n \rho_i \bigl\langle \xi_{*i}(t) \, \xi_{*i}(0) \bigr\rangle,$
and summing $t$ from $-\infty$ to $\infty$ yields,
\begin{equation}
  \sum_{k=1}^{n-1} \Gamma_k = \sum_{i=1}^n \rho_i \, G_i = \Gamma
  .
  \label{eq:Gammak_sum}
\end{equation}
This shows the equivalence of
Eq. \eqref{eq:EA_sbin1} and \eqref{eq:error_asym}
in the case of the single-bin updating scheme, with $\lambda_k \equiv 1$.
%In Appendix \ref{sec:Gamma_measure},
%we discuss a way of estimating $\Gamma_k$
%from an FDS simulation under a constant magnitude.
%

@@

We now determine the optimal value of $q(T)$
that minimizes the total error defined in Eq. \eqref{eq:error_tot}.
%
To model the initial error for Eq. \eqref{eq:error_res},
we will assume that at $t = 0$
the system has undergone a preliminary FDS simulation
under a constant updating magnitude, $a_0$,
for $T_0$ steps.
%
If the initial bias potential
before the preliminary run is zero,
one can show that
\begin{equation}
  \left\langle
    {\tilde v}_k^2(0)
  \right\rangle
  = \frac 1 2 \, a_0 \, \Gamma_k \, \lambda_k
  + \epsilon_k
  ,
  \label{eq:xt2_eql1}
\end{equation}
%
where
$\epsilon_k \equiv \tilde u_k^2 \, e^{-2\, \lambda_k \, a_0 \, T_0}$.
%
If $\lambda_k$ is not close to zero, and $T_0 \to \infty$
we can ignore the $\epsilon_k$ term, and
%
\begin{equation}
  \left\langle
    {\tilde v}_k^2(0)
  \right\rangle
  \approx
  \frac 1 2 \, a_0 \, \Gamma_k \, \lambda_k
  .
  \label{eq:xt2_eql}
\end{equation}
%
\note{
Consider a long period, $T_0$, under a fixed schedule,
$\alpha(t) = a_0$,
the residual error becomes negligible, and
the $k$th component of the asymptotic error
is given by
%
\begin{align*}
  \left\langle
    {\tilde v}_k^2(T_0)
  \right\rangle
  =
  \int_0^{T_0}
    \Gamma_k \, (\lambda_k \, a_0)^2 \,
      e^{ 2 \, \lambda_k \, a_0 \, (t - T_0) }
    \, dt
  \stackrel{ T_0 \to \infty }
  { =\joinrel=\joinrel=\joinrel= }
  \frac 1 2 \, \Gamma_k \, \lambda_k \, a_0
  .
\notag
%\label{eq:error_eql}
\end{align*}
}
%

>>>>

Free energy calculation\cite{frenkel, newman} is a central theme
in computational physics and chemistry
that can provide insight into an array of phenomena not easily studied
with regular simulation methods.
\note{``Regular simulation methods'' means MC and MD,
and they are more pertinent to ``computational physics and chemistry''
than ``traditional experiments.''}%
%
Given a system,
often the task is to compute
a distribution, $p(z)$,
along a collective variable, $z$, by sampling via either
Monte Carlo (MC)\cite{frenkel, newman, landau_binder}
or molecular dynamics (MD)\cite{frenkel, karplus2002}
simulations.
%
The distribution is related to
the dimensionless free energy, or potential of mean force (PMF),
along the collective variable, $z$,
as $-\ln p(z)$.
%
However, reliable estimation of the PMF is often difficult
when the collective variable is energetically restricted to local regions
of the complicated, multimodal free energy surface.
%
Thus,
to overcome these energetic barriers in sampling,
%and to capture the global shape of the free energy surface
an effective strategy is to introduce
a bias potential\cite{torrie1974, *torrie1977}
that offsets the PMF
such that the distribution along $z$ is
flat.\cite{mezei1987, berg1992, *lee1993,
wang2001, wang2001pre,
huber1994,
*laio2002, *laio2008, *barducci2011, *sutto2012}
%
\note{The flat-distribution sampling
allows one to more accurately capture the
global shape of the free energy surface
by improving the sampling around the rarely visited barrier region
and accelerating transitions among free energy basins.}

Many efficient flat-distribution-sampling (FDS) techniques
based on adaptive construction of the bias potential
have been introduced,
including the Wang-Landau (WL) algorithm\cite{
  wang2001, wang2001pre}
and metadynamics\cite{huber1994,
  *laio2002, *laio2008, *barducci2011, *sutto2012, micheletti2004}
among others\cite{yan2004, kim2006, *kim2007, kim2010, junghans2014,
  langfeld2012, pellegrini2014,
  maragliano2006, *abrams2008,
  zheng2010}.
%
These techniques regularly update the bias potential
to discourage future visits to previously sampled configurations
by incremental local elevation of the bias potential.
%
A key difference lies
in the updating window function,
which specifies
a neighborhood around the current value of
$z$ on the free energy surface
%where the updating should occur
as well as the relative updating magnitude.
%
The window function often
takes the form of a discrete
$\delta$-function (a boxcar function one histogram bin wide)
in the WL algorithm,
or that of a Gaussian
in metadynamics.\cite{junghans2014}
%
We will refer to the two updating schemes
as the single-bin
and Gaussian updating schemes, respectively.
%
While the former is popular for MC simulations,\cite{wang2001,
  wang2001pre, kim2006, *kim2007}
the latter is commonly used for MD simulations
as it provides a smoother bias potential
that can be readily differentiated
to derive the bias force.\cite{huber1994,
  *laio2002, *laio2008, *barducci2011, *sutto2012, junghans2014}
%
In both cases,
regular updates to the bias potential
disrupts the underlying equilibrium
sampling,\cite{zhou2005, morozov2007, zhou2008}
and the convergence of the bias potential
can often benefit from a gradually decreasing
updating magnitude of the bias potential.

The schedule of reducing the updating magnitude
affects the precision and accuracy
of the final bias potential,
and hence that of the PMF.\cite{
belardinelli2007, *belardinelli2007jcp, *belardinelli2008, *belardinelli2016,
liang2007,
morozov2007, morozov2009, zhou2008,
komura2012, *caparica2012, *caparica2014,
min2007,
barducci2008, dickson2011, dama2014, poulain2006}
%
While an overly slow reduction of the updating magnitude
is ineffective in restricting the variance of the bias potential,
a too rapid reduction often
converges the bias potential to a wrong value.
%
For the single-bin case, it is known that
the optimal schedule is given by the
inverse time\cite{
belardinelli2007, *belardinelli2007jcp, *belardinelli2008, *belardinelli2016,
liang2007,
morozov2007, morozov2009, zhou2008}
or
the inverse number of updating steps.

@@

We will then show by comparing the errors
of simulations under different updating schemes
that the single-bin updating scheme
is optimal for very long simulations.
%without a priori assumption of the smoothness of the PMF.
%
A similar but partial asymptotic optimality applies to
a class of bandpass updating schemes
that flatten only
a few long-range distribution modes.
%
Compared with the Gaussian updating scheme,
bandpass updating schemes have the advantage of allowing
simpler control of the updating magnitude
for optimal convergence.

@@

From Eqs.
\eqref{eq:vt_diffeq} and \eqref{eq:sh_ave},
we get % a set of decoupled equations
%
\begin{equation}
  \dot v_{*i}(t)
  %=
  %\alpha(t) \, \left[ \langle f_{*i}(t) \rangle + \xi_{*i}(t) \right]
  \approx
  -\alpha(t) \, \left[ v_{*i}(t) - \xi_{*i}(t) \right]
  .
  %\notag
  \label{eq:dxdt_singlebin}
\end{equation}
%
Without the random part, $\xi_{*i}(t)$,
the deviation of the bias potential, $v_{*i}(t)$,
would tend to $0$ at long times, reaching convergence.
%
The formal solution of Eq. \eqref{eq:dxdt_singlebin} is
%
\begin{equation}
  v_{*i}(T)
  =
  v_{*i}(0) \, e^{-q(T)}
  +
  \int_0^T
    \dot{\theta}\bigl( q(t) \bigr) \, \xi_{*i}(t) \, dt,
  \label{eq:xt_solution}
\end{equation}
%
where
%
$q(t) \equiv \int_0^t \alpha(t') \, dt'$,
%
and
%
\begin{align}
  \theta(q')
  &\equiv
  e^{q' - q(T)}.
  \label{eq:theta_def}
\end{align}

@@

From Eqs. \eqref{eq:error_def} and \eqref{eq:xt_solution},
we can write the total error at the end of period $T$
as a sum of a residual error, $E_R$ (from the decay of the initial error)
and an asymptotic error, $E_A$:
%
\begin{align}
  \Err(T)
  =
  \Err_R(T) + \Err_A(T)
  ,
  \label{eq:error_tot}
\end{align}
%
where
\begin{align}
  \Err_R(T)
  &= \Err(0) \, e^{-2 \, q(T)}
  ,
  \label{eq:ER_sbin}
  \\
  \Err_A(T)
  &=
  \sum_{i=1}^n \rho_i
    \int_0^T\!\!\!\int_0^T
    \!\!
    \langle
      \xi_{*i}(t) \xi_{*i}(t')
    \rangle
    \dot \theta\bigl( q(t) \bigr)
    \dot \theta\bigl( q(t') \bigr)
    \, dt dt'
  \notag \\
  &= \Gamma \, \int_0^T \dot \theta^2\bigl( q(t) \bigr) \, dt
  ,
  \label{eq:EA_sbin}
\end{align}
with
%$\Gamma \equiv \sum_{i=1}^n \rho_i \, G_i$.
$\Gamma \equiv \sum_{i=1}^n \rho_i \, G_i
\approx \sum_{i=1}^n \rho_i \int_{-\infty}^{\infty} \langle \xi_{*i}(t) \xi_{*i}(0) \rangle \, dt$.
%\begin{equation}
%  \Gamma \equiv \sum_{i=1}^n \rho_i \, G_i
%  =
%  \sum_{i=1}^n \rho_i \int_0^T
%    \langle \xi_{*i}(t) \xi_{*i}(0) \rangle \, dt
%  .
%  \notag%\label{eq:Gamma_from_G}
%\end{equation}

The error depends implicitly on the curve, $\alpha(t)$,
or, equivalently, the curve, $q(t)$.
%via $v_{*i}(T)$.
%
%If $\alpha(t)$ is a constant, $a_0$,
%so that $q(t) = a_0 \, t$,
%then after a long period,
%the error approach an equilibrium value
%%
%\begin{equation}
%  \Err_\mathrm{equil.}
%  = \Err_A = \Gamma \int_0^\infty e^{-2\,q'} \, a_0 \, d q'
%  = \frac{1}{2} a_0 \Gamma
%  .
%  \label{eq:error_equil_sbin}
%\end{equation}
%
Our aim is to find the $q(t)$
that minimizes this error.
%
It is convenient to
vary the curve, $q(t)$ ($0 < t < T$),
under a fixed value of $q(T)$,
since it allows us to focus on
the asymptotic error, $E_A(T)$.
%
The contribution of the residual error
will be considered later
by adjusting the value of $q(T)$.
%
Note that the asymptotic error
has the form of an action of a free particle
at $\theta\bigl( q(t) \bigr)$,
with the endpoints fixed at
$\theta\bigl( q(0) \bigr)  = e^{- q(T)}$
and
$\theta\bigl( q(T) \bigr) = 1$.
%
Thus, the velocity should be a constant
%Since the residual error
%is fixed during the process,
%the asymptotic error,
%resembling the action of a free particle,
%demands a constant velocity
%
\begin{equation}
  \frac{d}{dt} \theta\bigl( q(t) \bigr) = c
  .
\label{eq:dthetadt_const}
\end{equation}
%
Then, by using Eq. \eqref{eq:theta_def}, we have
%
\begin{equation}
  e^{ q(t) - q(T) }
  =
  c \, (t + t_0)
  =
  \frac{ t + t_0 }
       { T + t_0 }
  ,
\label{eq:expqt}
\end{equation}
%
where $t_0$ and $c$ are two constants,
and the latter has been determined
from the $t = T$ case
in the second step.
%
Taking the logarithm,
and differentiating both sides
with respect to $t$,
we get the inverse-time schedule
%
\begin{equation}
  \alpha(t) = \frac{ 1 }{ t + t_0 }
  ,
\label{eq:alpha_invt1}
\end{equation}
%
with Eq. \eqref{eq:alpha_invt} being
the special case of $t_0 = 0$.
%

@@

Equation \eqref{eq:vt_diffeq_mbin} shows that
in the absence of the fluctuating term, $\tilde \xi_k(t)$,
the deviation of the bias potential would tend to zero
at a rate of $\alpha(t) \, \lambda_k$.
%
Thus, the updating scheme is convergent
only if no eigenvalue, $\lambda_k$, is negative
for $k > 0$.
%
%With a set of heterogeneous eigenvalues,
%the errors of the eigenmodes are reduced at different rates.
%Thus, the optimization problem of the final error
%becomes more complicated and it often
%involves a compromise among eigenmodes.
%
The formal solution is given by
\begin{equation}
  \tilde v_k(T)
  =
  \tilde v_k(0) \, e^{-\lambda_k \, q(T)}
  +
  \int_0^T
    \dot{\theta}_k\bigl( q(t) \bigr) \, \tilde\xi_k(t) \, dt
  ,
  \label{eq:xt_solution}
\end{equation}
where
\begin{equation}
  \theta_k(q') \equiv e^{\lambda_k \, [q' - q(T)]}
  .
  \label{eq:uk_def}
\end{equation}

@@

By Eq. \eqref{eq:parseval},
we can rewrite the error defined in Eq. \eqref{eq:error_def} as
a sum of the errors in individual modes\footnote{If
the error weighting distribution differs from $\rho_i$
in Eq. \eqref{eq:error_def},
the error would contain quadratic cross terms among different eigenmodes.
%
While the solution for the optimal schedule retains the same form,
the mass function defined in Eq. \eqref{eq:mass_func}
will include the corresponding cross terms.
}
%$\Err(T) = \sum_{k = 1}^{n - 1} \left\langle {\tilde v}_k^2(T) \right\rangle$.
%
\begin{align}
  \Err(T)
  =
  \sum_{k = 1}^{n - 1}
    \left\langle
      {\tilde v}_k^2(T)
    \right\rangle
  .
  \notag
\end{align}
%
As in Eq. \eqref{eq:error_tot},
we can decompose the error as the sum
of the residual error and the asymptotic error, with
\begin{align}
  \Err_R(T)
  &=
  \sum_{k = 1}^{n-1}
    \left\langle
      {\tilde v}_k^2(0)
    \right\rangle \,
    e^{ - 2 \, \lambda_k  \, q(T) }
  ,
  \label{eq:error_res}
  \\
  \Err_A(T)
  &=
  \sum_{k=1}^{n-1}
  \int_0^T\!\!\!\int_0^T
  \!\!
  \langle
    \tilde\xi_k(t)
    \tilde\xi_k(t')
  \rangle
  \theta_k\bigl(q(t)\bigr)
  \theta_k\bigl(q(t')\bigr)
  \, dtdt'
  \notag \\
  &=
  \int_0^T
  \sum_{k = 1}^{n-1}
  \Gamma_k \, \dot \theta_k^2\bigl( q(t) \bigr) \, dt
  ,
  \label{eq:error_asym}
\end{align}
%
where
%
%$\Gamma_k$ is computed from
%the autocorrelation function
%of the histogram fluctuation,
%$\bigl\langle {\tilde \xi}_k(0)
%\, {\tilde \xi}_k(t) \bigr\rangle$,
%as:
%
\begin{equation}
  \Gamma_k
  \approx
  \int_{-\infty}^\infty
  \bigl\langle
    \tilde\xi_k(t) \, \tilde\xi_k(0)
  \bigr\rangle
  \, dt
  \to
  \sum_{t = -\infty}^\infty
  \bigl\langle
    \tilde\xi_k(t) \, \tilde\xi_k(0)
  \bigr\rangle
  ,
  \label{eq:Gamma_sum}
\end{equation}
%
with the second expression being
the counterpart in the discrete-time setting.
%
Equation \eqref{eq:Gamma_sum} represents
a first approximation for modeling
a general fluctuation, ${\tilde \xi}_k(t)$,
as an equivalent white noise.
In this way,
the underlying sampling process
affects the error only through the
few numbers, $\Gamma_k$'s.
%
Note that by Eq. \eqref{eq:parseval}, we have
$\sum_{k=1}^{n-1} \bigl\langle {\tilde \xi}_k(t) \, {\tilde \xi}_k(0) \bigr\rangle
= \sum_{i=1}^n \rho_i \bigl\langle \xi_{*i}(t) \, \xi_{*i}(0) \bigr\rangle,$
and summing $t$ from $-\infty$ to $\infty$ yields,
\begin{equation}
  \sum_{k=1}^{n-1} \Gamma_k = \sum_{i=1}^n \rho_i \, G_i = \Gamma
  .
  \label{eq:Gammak_sum}
\end{equation}
This shows the equivalence of
Eq. \eqref{eq:EA_sbin1} and \eqref{eq:error_asym}
in the case of the single-bin updating scheme, with $\lambda_k \equiv 1$.
%In Appendix \ref{sec:Gamma_measure},
%we discuss a way of estimating $\Gamma_k$
%from an FDS simulation under a constant magnitude.
%

@@

We now determine the optimal value of $q(T)$
that minimizes the total error defined in Eq. \eqref{eq:error_tot}.
%
To model the initial error for Eq. \eqref{eq:error_res},
we will assume that at $t = 0$
the system has undergone a preliminary FDS simulation
under a constant updating magnitude, $a_0$,
for $T_0$ steps.
%
If the initial bias potential
before the preliminary run is zero,
we have, from Eq. \eqref{eq:xt_solution} [with $T\to T_0, q(t) \to a_0 \, t$],
%
\begin{align}
  \left\langle {\tilde v}_k(0) \right\rangle
  &=
  \tilde u_k \, e^{-\lambda_k \, a_0 \, T_0}
  ,
  \notag \\
  \operatorname{var} \tilde v_k(0)
  &=
  \frac 1 2 \, \Gamma_k \, \lambda_k \, a_0
  \left( 1 - e^{-2 \, \lambda_k \, a_0 \, T_0} \right)
  ,
  \label{eq:xtvar}
  \\
  \left\langle \tilde v_k^2(0) \right\rangle
  &=
  %\operatorname{var} \tilde v_k(0) + \zeta_k^2
  %=
  \frac 1 2 \, \Gamma_k \, \lambda_k \, a_0
  +
  \epsilon_k \, e^{-2 \, \lambda_k \, a_0 \, T_0}
  .
  \label{eq:xt2_eql}
\end{align}
%
where
$\epsilon_k \equiv \left(\tilde u_k^2 - \frac{1}{2} \Gamma_k \, \lambda_k \, a_0\right) \, e^{-2\,\lambda_k a_0 T_0}$.
